---
title: "Group Project"
author: "Lucien Patchett"
date: "2025-10-30"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Discussion
Fibromyalgia is a medical condition that presents through symptoms “including fatigue, sleep disturbances, cognitive dysfunction, and depressive episodes… chronic fatigue syndrome, irritable bowel syndrome (IBS), irritable bladder syndrome or interstitial cystitis, and temporomandibular disorder (TMD)” (Clauw, 2009). Patients with fibromyalgia often face daily challenges as they struggle to live a normal life due to the limited visual indicators that can be perceived by those around them. With this condition occurring in 2-8% of the global population (Clauw, 2014) it has become crucial that we find ways to mitigate the effects on patients.
There is significant research on the effects of music as a non-pharmaceutical intervention to reduce pain levels in patients, however in this paper we will analyze the effects of non-musical (pink) noise as it compares to music to determine if there is significance in music as a treatment option for fibromyalgia.

#Question
Can music and/or pink noise be used as an analgesic in patients with fibromyalgia?

#Hypothesis
Music and pink noise will lead to a decrease in nociception in patients with a diagnosis of fibromyalgia.

#Prediction
Participants who present with a diagnosis fibromyalgia will find a greater decrease in nociception when listening to music when compared to "pink" noise.

#pic and pim boxplot
``` {r code chunk pim/pic}
library(ggplot2)
library(tidyr)
library(dplyr)

data <- read.csv("dataset.csv")

# Pivot pim1 and pim2 into long format
data_long_pim <- data %>%
  pivot_longer(cols = c(pic1,pic2,pim1,pim2), names_to = "treatment", values_to = "value")

# Make boxplot
ggplot(data_long_pim, aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot(alpha = 0.7, width = 0.6) +
  theme_classic() +
  scale_fill_manual(values = c("pic1" = "Red", "pic2" = "Green", "pim1" = "Orange", "pim2" = "Black")) +
  ylab("Pain Level") +
  xlab("") +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    legend.position = "none"
  )+
    scale_x_discrete(labels = c("pic1" = "Before Noise", "pic2" = "After Music", "pim1" = "Before Music", "pim2" = "After Music"))

```


#Pic1 and Pic2 code chunk
``` {r code pic chunks}

library(ggplot2)
library(tidyr)
library(dplyr)

data <- read.csv("dataset.csv")

# Pivot pic1 and pic2 into long format
data_long <- data %>%
  pivot_longer(cols = c(pic1,pic2), names_to = "treatment", values_to = "value")

# Make boxplot
ggplot(data_long, aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot(alpha = 0.7, width = 0.6) +
  theme_classic() +
  scale_fill_manual(values = c("pic1" = "Red", "pic2" = "Green")) +
  ylab("Pain Level") +
  xlab("") +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    legend.position = "none"
  )+
  scale_x_discrete(labels = c("pic1" = "Before", "pic2" = "After"))

```

#Pim1 and Pim2 code chunk
``` {r code chunk pim}
library(ggplot2)
library(tidyr)
library(dplyr)

data <- read.csv("dataset.csv")

# Pivot pim1 and pim2 into long format
data_long_pim <- data %>%
  pivot_longer(cols = c(pim1,pim2), names_to = "treatment", values_to = "value")

# Make boxplot
ggplot(data_long_pim, aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot(alpha = 0.7, width = 0.6) +
  theme_classic() +
  scale_fill_manual(values = c("pim1" = "Orange", "pim2" = "Black")) +
  ylab("Pain Level") +
  xlab("") +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    legend.position = "none"
  )+
    scale_x_discrete(labels = c("pim1" = "Before", "pim2" = "After"))

```

``` {r code chunk pim vs pic}
library(ggplot2)
library(tidyr)
library(dplyr)

data <- read.csv("dataset.csv")

# Pivot pic2 and pim2 into long format
data_long_pim <- data %>%
  pivot_longer(cols = c(pic2,pim2), names_to = "treatment", values_to = "value")

# Make boxplot
ggplot(data_long_pim, aes(x = treatment, y = value, fill = treatment)) +
  geom_boxplot(alpha = 0.7, width = 0.6) +
  theme_classic() +
  scale_fill_manual(values = c("pic2" = "Green", "pim2" = "Black")) +
  ylab("Pain Level") +
  xlab("") +
  theme(
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    legend.position = "none"
  )+
    scale_x_discrete(labels = c("pic2" = "After Noise", "pim2" = "After Music"))

```


#Works Cited
OpenAI. (2025). ChatGPT (GPT-5) [Large language model]. OpenAI. https://chat.openai.com/

Clauw, D. J. (2009). Fibromyalgia: An Overview. The American Journal of Medicine, 122(12), S3–S13. https://doi.org/10.1016/j.amjmed.2009.09.006

Clauw, D. J. (2014). Fibromyalgia. JAMA, 311(15), 1547. https://doi.org/10.1001/jama.2014.3266